---
title: "Exploration of Diamond Pricing"
author: "Paul Foley"
output: html_document
---


Explore the diamond data provided by ggplot 2, and specifically to how diamond price is affected by its various features:

* Cut
* Carat
* Color
* Clarity
* Volume
* Depth
* Table

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_Libraries}
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(lattice)
library(MASS)
library(car)
library(reshape2)
library(dplyr)
library(gridExtra)
library(RColorBrewer)
library(bitops)
library(RCurl)
library(knitr)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
data(diamonds)
```

## Diamond Data Summary

```{r echo=FALSE, message=FALSE, warning=FALSE, Summary}
summary(diamonds)
?diamonds
```

## Price Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, Summary-Price}
summary(diamonds$price)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Histogram-Price(log10)}
plot1 <- ggplot(data = diamonds,
                aes(x = price)) +
  geom_histogram(binwidth = 500,
                 fill = 'blue') +
  scale_x_continuous(limits = c(0, 20000),
                     breaks = seq(0, 20000, 5000)) +
  xlab('Price') +
  ylab('Number of Diamonds') +
  ggtitle('Histogram - Price')

plot2 <- ggplot(data = diamonds,
                 aes(x = price)) +
  geom_histogram(binwidth = .01,
                 fill = 'orange') +
  ggtitle('Price (log10)') +
  scale_x_log10() +
  ggtitle('Histogram - Price (Log10)')

grid.arrange(plot1,plot2, ncol = 2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Histogram-Price-Cheap}
ggplot(data = diamonds,
       aes(x = price)) +
  geom_histogram(binwidth = 10,
                 fill = 'blue') +
  scale_x_continuous(limits = c(0, 2000),
                     breaks = seq(0, 2000, 200)) +
  xlab('Price') +
  ylab('Number of Diamonds') +
  ggtitle("Diamonds under $2000")
```

## Price by Cut Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, Price-Cut}
by(diamonds$price, diamonds$cut, summary)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Boxplot-Price-Cut}
ggplot(data = diamonds,
       aes(x = cut,
           y = price)) +
  geom_boxplot(fill = 'green') +
  coord_cartesian(ylim = c(0, 10000)) +
  ggtitle("Boxplots - Price Vs. Cut")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Histogram-Price-Cut}
ggplot(data = diamonds,
       aes(x = price)) +
  geom_histogram(binwidth = 500,
                 fill = 'blue') +
  scale_x_continuous(limits = c(0, 20000),
                     breaks = seq(0, 20000, 5000)) +
  facet_wrap(~cut,
             scales='free_y') +
  xlab('Price') +
  ylab('Number of Diamonds') +
  ggtitle('Histograms - Price Group By Cut')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Histogram-Price-Carat-Cut}
ggplot(data = diamonds,
       aes(x = price/carat)) +
  geom_histogram(binwidth = .05,
                 fill = 'blue') +
  scale_x_log10() +
  facet_wrap(~cut,
             scales='free_y') +
  xlab('Price Per Carat') +
  ylab('Number of Diamonds') +
  ggtitle('Histograms - Price per Carat Group By Cut')
```

## Price By Carat Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, Frequency-Carat}
ggplot(data = diamonds,
       aes(x = carat)) +
  geom_freqpoly(binwidth = .01) +
  scale_x_continuous(limits = c(0, 3),
                     breaks = seq(0, 3, 1)) +
  xlab('Weight') +
  ylab('Number of Diamonds') +
  ggtitle('Frequency Polygon - Carat')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Price-Carat}
summary(diamonds$price/diamonds$carat)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price-Carat}
ggplot(data = diamonds,
       aes(x = carat,
           y = price)) +
  geom_point(alpha = .1,
             color = 'green') +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99)) +
  geom_smooth(method = 'lm',
              color = 'red') +
  ggtitle('Scatterplot - Price Vs. Carat')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Carat_Price-log10}
ggplot(data = diamonds,
       aes(carat,
           price)) +
  geom_point(alpha = .1,
             color = 'green') +
  scale_y_continuous(trans = log10_trans()) +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  geom_smooth(method = 'lm',
              color = 'red') +
  ggtitle('Scatterplot - Price(log10) Vs. Carat')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Carat_Price_Transform}
cuberoot_trans = function() trans_new('cuberoot',
                                      transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
ggplot(data = diamonds,
       aes(carat, 
           price)) +
  geom_point(alpha = .1,
             color = 'green') +
  scale_x_continuous(trans = cuberoot_trans(),
                     limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) +
  scale_y_continuous(trans = log10_trans(),
                     limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  geom_smooth(method = 'lm',
              color = 'red') +
  ggtitle('Scatterplot - Price (log10) Vs. Carat(Cube-Root)')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price_Carat_Clarity}
ggplot(data = diamonds,
       aes(x = carat, y = price)) +
  geom_point(alpha = 0.5,
             size = 1, 
             position = 'jitter',
             aes(color = clarity)) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity',
                                         reverse = TRUE,
                                         override.aes = list(alpha = 1, 
                                                             size = 2))) +
  scale_x_continuous(trans = cuberoot_trans(),
                     limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) +
  scale_y_continuous(trans = log10_trans(),
                     limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) Vs. Carat (Cube Root) \n Color By Clarity')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price_Carat_Cut}
ggplot(data = diamonds, 
       aes(x = carat,
           y = price,
           color = cut)) + 
  geom_point(alpha = 0.5,
             size = 1,
             position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut',
                                          reverse = T,
                                          override.aes = list(alpha = 1,
                                                              size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(),
                     limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(),
                     limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) Vs. Carat (Cube-Root) \n Color By Cut')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price_Carat_Color}
ggplot(data = diamonds,
       aes(x = carat,
           y = price,
           color = color)) + 
  geom_point(alpha = 0.5,
             size = 1,
             position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color',
                                          override.aes = list(alpha = 1,
                                                              size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(),
                     limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(),
                     limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) Vs. Carat (Cube-Root) \n Color By Color')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price-Carat-Colored}
diamonds$price_per_carat <- diamonds$price / diamonds$carat

ggplot(data = diamonds,
       aes(x = cut,
           y = price_per_carat)) +
  geom_jitter(alpha = .5,
              aes(color = color)) +
  facet_wrap(~clarity) +
  scale_color_brewer(type = 'div') +
  ggtitle('Scatterplots - Price Per Carat Vs. Cut \n Group By Clarity and Color by Color')
```

## Price By Color Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, Price-Color}
by(diamonds$price, diamonds$color, summary)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, IQR-Price-Color}
ggplot(data = diamonds,
       aes(x = color,
           y = price)) +
  geom_boxplot(fill = 'green') +
  coord_cartesian(ylim = c(0, 10000)) +
  ggtitle('Interquartile Range (IQR) - Price Vs. Color')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Boxplots-Price-Carat}
ggplot(data = diamonds,
       aes(x = color,
           y = price/carat)) +
  geom_boxplot(fill = 'green') +
  coord_cartesian(ylim = c(0, 7500)) +
  ggtitle('Boxplots - Price Per Carat Vs. Color')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Histogram-Facet-Color}
ggplot(data = diamonds,
       aes(x = price)) +
  geom_histogram(binwidth = .05,
                 aes(fill = cut)) +
  facet_wrap(~color) +
  scale_fill_brewer(palette = "Spectral") +
  scale_x_log10() +
  ggtitle('Histogram - Price \n Group By Color, Color By Cut')
```

## Price By Clarity Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, Barchart-Mean-Price}
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n()) %>%
  arrange(clarity)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, 
                                    mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color,
                                 mean_price = mean(price))

diamond_clarity_plot <- ggplot(data = diamonds_mp_by_clarity, 
                               aes(x = clarity, y = mean_price)) +
  geom_bar(stat="identity",
           fill = 'green') +
  ggtitle('Bar Chart - Mean Price Vs. Clarity')

diamond_color_plot <- ggplot(data = diamonds_mp_by_color,
                             aes(x = color, y = mean_price)) +
  geom_bar(stat="identity",
           fill = 'green') +
  ggtitle('Bar Chart - Mean Price Vs. Color')

grid.arrange(diamond_clarity_plot, diamond_color_plot)
```

## Price By X, Y, Z Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlations-Price}
with(diamonds, cor.test(price, x, method = 'pearson'))
with(diamonds, cor.test(price, y, method = 'pearson'))
with(diamonds, cor.test(price, z, method = 'pearson'))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price-X}
ggplot(data = diamonds,
       aes(x = x, y = price)) +
  geom_point(alpha = .1,
             color = 'green')+
  xlim(0, quantile(diamonds$x, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99)) +
  geom_smooth(method = 'lm',
              color = 'red') +
  ggtitle('Scatterplot - Price Vs. X')
```

### Correlation - Price Vs. Volume

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlation-Price-Volume}
diamonds$volume <- diamonds$x * diamonds$y + diamonds$z
volume_df <- subset(diamonds, volume > 0 & volume < 800)

cor.test(volume_df$price, volume_df$volume, method = "pearson")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price-Volume}
ggplot(data = diamonds,
       aes(x = volume,
           y = price)) +
  geom_point(alpha = .1,
             color = 'green') +
  xlim(0, quantile(diamonds$volume, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99)) +
  geom_smooth(method = 'lm',
              color = 'red') +
  ggtitle('Scatterplot - Price Vs. Volume')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price-Volume-2}
ggplot(data = volume_df,
       aes(x = volume,
           y = price)) +
  geom_point(alpha = .1,
             color = 'green') +
  xlim(0, quantile(volume_df$volume, 0.99)) +
  ylim(0, quantile(volume_df$price, 0.9)) +
  geom_smooth(method = 'lm',
              color = 'red') +
  ggtitle('Scatterplot - Price Vs. Volume Filtered')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price-Volume-Clarity}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(data = diamonds,
       aes(x = volume, 
           y = price)) +
  geom_point(aes(color = clarity)) +
  scale_color_brewer(type = 'div') +
  xlim(0, quantile(diamonds$volume, 0.99)) +
  scale_y_log10() +
  ggtitle('Scatterplot - Price (log10) Vs. Volume \n Color By Clarity')
```

## Price By Depth Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlation-Price-Depth}
with(diamonds, cor.test(price, depth, method = 'pearson'))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatterplot-Price-Depth}
ggplot(data = diamonds, aes(x = depth,
                            y = price)) +
  geom_point(alpha = .05,
             color = 'green') +
  scale_x_continuous(limits = c(56, 66),
                     breaks = seq(58, 64, 2)) +
  ggtitle('Scatterplot - Price Vs. Depth')
```

## Price by Table Exploration

```{r echo=FALSE, message=FALSE, warning=FALSE, Table-Price-Cut}
ggplot(data = diamonds,
       aes(x = table,
           y = price)) +
  geom_point(aes(color = cut)) +
  scale_color_brewer(type = 'qual') +
  xlim(50, 80) + 
  ggtitle('Scatterplot - Price Vs. Table \n Color By Cut')
```

## Building the Linear Model for Price

```{r echo=FALSE, message=FALSE, warning=FALSE, Linear_Model}
m1 <- lm(data = diamonds,
         I(log(price)) ~ I(carat^(1/3)))
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)

mtable(m1, m2, m3, m4, m5, sdigits = 3)
```